### 2022-03-29 paper/Chapters/Hardware.tex
Tweaks and updates to the hardware sections
Now a nice big sprint can begin. Just need the time

@@

When starting to choose parts for the hardware, there are many difficult questions to face.
First and foremost, the hardware had to be built with the time and resources available for this project.
Specifically, building a Single Board Computer\todoadvisor{introduce acronyms the first time they're used}\todo{This was introduced in 4.1.1, should it also be done here?} from scratch would be completely infeasible, but the goal was also not to simply purchase an off the shelf SBC that was not tailored to the task at hand such as a traditional Raspberry Pi.
Instead, a better suited solution would be to utilize a \emph{Raspberry Pi Compute Module 4 (CM4)} which has \enquote{The power of Raspberry Pi 4 in a compact form factor for deeply embedded applications} \cite{rpi_cm4}.
Boasting an incredibly small size and streamlined I/O, there only way to communicate with the Compute Module was through two tiny mezzanine connectors as shown in Figure \ref{fig:rpi_cm4_mezzanine}.
With no other pin outs for anything such as power, HDMI, or USB, everything had to be built on a custom Printed Circuit Board (PCB).

@@

This PCB would have to be designed so that the CM4 could be plugged into the board using the mezzanine connectors and rely on it for all inputs and outputs.
This, however, is only the beginning of what is required to develop the product.
Since the project would not use an off the shelf SBC like a regular Raspberry PI, the project would require only designing the PCB but also picking out all the components that would go onto the PCB, such as the USB ports themselves, the USB driver chip, and related components needed for operation.
Because the CM4 is designed to be as low-powered and compact as possible, some features that would usually be expected of such a device, like the USB ports, are disabled by default and must instead be built by the board designer.
While this helps greatly by removing extraneous components and features that aren't needed for this project, it also provides me with a \todo{does this sound sarcastic?}great opportunity to learn exactly what parts are being used in the hardware.

An invaluable resource while designing the PCB was the Official I/O board for the CM4 \cite{rpi_cm4_io}.
Though the PCB board would still have to be custom built and components would need to be replaced due to the current ongoing chip shortage, it provides a great foundation instead of starting from scratch.
Another massive benefit from working within the Raspberry Pi Ecosystem is their \todo{(re)move?}incredible documentation and community surrounding the boards.
Around the time this project was being started, the CM4 was just starting to get into the hands of makers and developers, and only a couple custom boards had actually been made by anyone other than Raspberry Pi themselves.
Even though this leaves the community with less of a bulk of knowledge to work with, it has provided a sort of comradery with everyone helping each other learn the new Compute Module together.
Members of the CM4 engineering team would even assist those on the forums to help them pick parts and design their own boards.

@@

\renewcommand*{\arraystretch}{1.5} % inner table padding
\begin{longtable}{|>{\raggedright\arraybackslash}p{0.2\linewidth}|>{\raggedright\arraybackslash}p{0.33\linewidth}|>{\raggedright\arraybackslash}p{0.46\linewidth}|}
  \hline
  \bfseries Purpose & \bfseries Manufacturer's Number &\bfseries Reason \\% Columns
  \hline
  USB-C Power in & \emph{USB4125-GF-A} by \enquote{GCT} & Most projects of this scale are powered by micro usb, but with USB-C gaining popularity, this component was picked for its low price and exclusive ability to transfer power without transfering data. \\
  \hline
  Ethernet Port & \emph{ARJM11D7-502-AB-EW2} by \enquote{Abracon LLC} & One of the few in-stock ethernet ports that supports \emph{1000 BASE-T} usage. The CM4 recommends the \emph{MagJack-A70-112-331N126} produced by \enquote{LINK-PP}, but that part is out of stock for at least another year as of writing. \\
  \hline
  USB Port & \emph{SS-52100-001} by \enquote{Stewart Connector} & Single flat port that supports USB 2.0. \\
  \hline
  HDMI Port & \emph{SS-53000-001} by \enquote{Stewart Connector} & Single flat port that supports HDMI 1.4. \\
  \hline
  Flat Flexible Cable Connector & \emph{FFC3B07-20-T} by \enquote{GCT} & A generic Flat Flexible Cable port that can be connected to anything. In this case, one is connected to the build-in screen's HDMI port, and another is connected to the screen's touch/power USB port. \\
  \hline
  2-Pin Through Hole Header & \emph{PREC002SAAN-RC} by \enquote{Sullins Connector Solutions} & A generic 2-pin header than can be used to bypass the power circuit as discussed in Section \ref{sec:Prototyping}. \\
  \hline
  HDMI Power Regulator & \emph{RT9742ENGJ5} by \enquote{Richtek USA Inc.} & The CM4 recommends the \emph{RT9742SNGV} from the same manufacturer, but it is out of stock. This part has slight differences that call for tweaks in the final layout. \\
  \hline
  SD-Card Power Regulator & \emph{RT9742VGJ5} by \enquote{Richtek USA Inc.} & The CM4 recommends the \emph{RT9742GGJ5F} from the same manufacturer, but it is out of stock. This part has slight differences that call for tweaks in the final layout. \\
  \hline
  Battery Voltage Regulator and Boost Converter & \emph{TPS61090RSAR} by \enquote{TPS61090RSAR} & Used to balance power usage from the USB-C port or the attached battery, favoring the external power when available. \\
  \hline
  Li-Po Battery Charge Management Controller & \emph{MCP73871-2CCI\_ML} by \enquote{Microchip Technology} & Used to charge the attached battery when attached to external power through the USB-C port. \\
  \hline
  4 Port USB Hub Controller & \emph{USB2504A-JT} by \enquote{Microchip Technology} & Used to control the four USB 2.0 ports and send the data to the CM4. The CM4 recommends the \emph{USB2514B-I/M2} by the same manufacturer, however it and all the parts in it's family are out of stock with at least a 52-week back order. Out of all the components, this has the highest potential to be problematic.\\
  \hline
  \caption[Selected PCB Components]{A non-exhaustive list of components that were chosen for this project.}
  \label{tab:SelectedComponents}
\end{longtable}
\todo[inline]{Should the above table have pictures? If so, where? Since the table is already pretty dense}

@@

Instead, the first board was built out in a piecewise fashion, allowing each section of the board a way to fail without interrupting the rest of the board.
This allows each function of the board to be tested individually and in isolation so that changes could be made incrementally without needing to worry about their integration with the rest of the board.
\todo{convert first person to third person}For example, in order to build a system that would allow the board to be powered via USB or an internal battery, as well as charge the battery simultaneously, I would need to build a complex circuit consisting of parts neither I nor anyone I knew had used before.
Since power is obviously a critical component of the board, A secondary way of feeding power directly the board was built to bypass all that circuitry in case it fails.
In doing so, the board can still be powered and it's other features can be tested while a fix for that portion of the circuitry is developed for the next iteration.

@@

Finally, everything must connect to the CM4.
Figure \ref{fig:CM4Circuit} shows the the two mezzanine connectors the CM4 uses to communicate with the rest of the PCB (schematic in Figure \ref{fig:pcb_schematic_cm4}).
Remember that there is only 0.04mm of space in between each of the pins, leaving very little space to route the traces where they need to go.
The first thing to notice is the large stretch of traces on the back side of the board coming from the HDMI ports.
Because most of the traces come in positive and negative pairs, called differential pairs, the two tracks must be roughly the same length as each other.
This is because as data is transmitted across the pair of tracks, the two signals need to arrive at their destination at the same time, a time that is determined by the physical distance the signals need to cover.
Since the speed of electricity is constant, the traces may need to wind back and forth so that the lengths positive and negative tracks are equal.
Also note the cutout area at the top of the CM4, whose perimeter is stitched by vias; this is where the CM4 houses it's bluetooth and wifi antennas.

\begin{figure}[t]
  \centering
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=.5\linewidth]{Figures/kicad/close-ups/cm4-front}
    \caption{The front side of the PCB}
    \label{fig:CM4CircuitFront}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=.5\linewidth]{Figures/kicad/close-ups/cm4-back}
    \caption{The back side of the PCB}
    \label{fig:CM4CircuitBack}
  \end{subfigure}
  \caption{The circuitry surrounding the CM4's two mezzanine connectors}
  \label{fig:CM4Circuit}
\end{figure}

>>>>

There are a lot of factors that go into designing and building an electrical system that fits the needs for this project.
This process begins by outlining the requirements in Section \ref{sec:HardwareRequirements}, followed by describing the process and methodology of choosing parts in Section \ref{sec:ChoosingParts}.
Then, the feasibility of prototyping such as system is discussed in Section \ref{sec:Prototyping}, and finally the system is designed and manufactured in Sections \ref{sec:DesigningThePCB} and \ref{sec:ManufacturingThePCB} respectively.

@@

When starting to choose parts for the hardware, there are many difficult questions to face.
First and foremost, the hardware had to be built with the time and resources available for this project.
Specifically, building a Single Board Computer\todoadvisor{introduce acronyms the first time they're used}\todoquestion{This was introduced in 4.1.1, should it also be done here?} from scratch would be completely infeasible, but the goal was also not to simply purchase an off the shelf SBC that was not tailored to the task at hand such as a traditional Raspberry Pi.
Instead, a better suited solution would be to utilize a \emph{Raspberry Pi Compute Module 4 (CM4)} which has \enquote{The power of Raspberry Pi 4 in a compact form factor for deeply embedded applications} \cite{rpi_cm4}.
Boasting an incredibly small size and streamlined I/O, there only way to communicate with the Compute Module was through two tiny mezzanine connectors as shown in Figure \ref{fig:rpi_cm4_mezzanine}.
With no other pin outs for anything such as power, HDMI, or USB, everything had to be built on a custom Printed Circuit Board (PCB).

@@

This PCB would have to be designed so that the CM4 could be plugged into the board using the mezzanine connectors and rely on it for all inputs and outputs.
This, however, is only the beginning of what is required to develop the product.
Since the project would not use an off the shelf SBC like a regular Raspberry PI, the project would require only designing the PCB but also picking out all the components that would go onto the PCB, such as the USB ports themselves, the USB driver chip, and related components needed for operation.
Because the CM4 is designed to be as low-powered and compact as possible, some features that would usually be expected of such a device, like the USB ports, are disabled by default and must instead be built by the board designer.
While this helps greatly by removing extraneous components and features that aren't needed for this project, it also provides me with a \todoquestion{does this sound sarcastic?}great opportunity to learn exactly what parts are being used in the hardware.

An invaluable resource while designing the PCB was the Official I/O board for the CM4 \cite{rpi_cm4_io}.
Though the PCB board would still have to be custom built and components would need to be replaced due to the current ongoing chip shortage, it provides a great foundation instead of starting from scratch.
Another massive benefit from working within the Raspberry Pi Ecosystem is their \todoquestion{(re)move?}incredible documentation and community surrounding the boards.
Around the time this project was being started, the CM4 was just starting to get into the hands of makers and developers, and only a couple custom boards had actually been made by anyone other than Raspberry Pi themselves.
Even though this leaves the community with less of a bulk of knowledge to work with, it has provided a sort of comradery with everyone helping each other learn the new Compute Module together.
Members of the CM4 engineering team would even assist those on the forums to help them pick parts and design their own boards.

\todo{Add paragraph on DSI vs HDMI}

\todo{Add paragraph RTC}

@@

\renewcommand*{\arraystretch}{1.5} % inner table padding
\begin{longtable}{|>{\raggedright\arraybackslash}p{0.2\linewidth}|>{\raggedright\arraybackslash}p{0.33\linewidth}|>{\raggedright\arraybackslash}p{0.46\linewidth}|}
  \hline
  \bfseries Purpose & \bfseries Manufacturer's Number &\bfseries Reason \\% Columns
  \hline
  USB-C Power in & \emph{USB4125-GF-A} by \enquote{GCT} & Most projects of this scale are powered by micro usb, but with USB-C gaining popularity, this component was picked for its low price and exclusive ability to transfer power without transfering data. \\
  \hline
  Ethernet Port & \emph{ARJM11D7-502-AB-EW2} by \enquote{Abracon LLC} & One of the few in-stock ethernet ports that supports \emph{1000 BASE-T} usage. The CM4 recommends the \emph{MagJack-A70-112-331N126} produced by \enquote{LINK-PP}, but that part is out of stock for at least another year as of writing. \\
  \hline
  USB Port & \emph{SS-52100-001} by \enquote{Stewart Connector} & Single flat port that supports USB 2.0. \\
  \hline
  HDMI Port & \emph{SS-53000-001} by \enquote{Stewart Connector} & Single flat port that supports HDMI 1.4. \\
  \hline
  Flat Flexible Cable Connector & \emph{FFC3B07-20-T} by \enquote{GCT} & A generic Flat Flexible Cable port that can be connected to anything. In this case, one is connected to the build-in screen's HDMI port, and another is connected to the screen's touch/power USB port. \\
  \hline
  2-Pin Through Hole Header & \emph{PREC002SAAN-RC} by \enquote{Sullins Connector Solutions} & A generic 2-pin header than can be used to bypass the power circuit as discussed in Section \ref{sec:Prototyping}. \\
  \hline
  HDMI Power Regulator & \emph{RT9742ENGJ5} by \enquote{Richtek USA Inc.} & The CM4 recommends the \emph{RT9742SNGV} from the same manufacturer, but it is out of stock. This part has slight differences that call for tweaks in the final layout. \\
  \hline
  SD-Card Power Regulator & \emph{RT9742VGJ5} by \enquote{Richtek USA Inc.} & The CM4 recommends the \emph{RT9742GGJ5F} from the same manufacturer, but it is out of stock. This part has slight differences that call for tweaks in the final layout. \\
  \hline
  Battery Voltage Regulator and Boost Converter & \emph{TPS61090RSAR} by \enquote{TPS61090RSAR} & Used to balance power usage from the USB-C port or the attached battery, favoring the external power when available. \\
  \hline
  Li-Po Battery Charge Management Controller & \emph{MCP73871-2CCI\_ML} by \enquote{Microchip Technology} & Used to charge the attached battery when attached to external power through the USB-C port. \\
  \hline
  4 Port USB Hub Controller & \emph{USB2504A-JT} by \enquote{Microchip Technology} & Used to control the four USB 2.0 ports and send the data to the CM4. The CM4 recommends the \emph{USB2514B-I/M2} by the same manufacturer, however it and all the parts in it's family are out of stock with at least a 52-week back order. Out of all the components, this has the highest potential to be problematic.\\
  \hline
  \caption[Selected PCB Components]{A non-exhaustive list of components that were chosen for this project.}
  \label{tab:SelectedComponents}
\end{longtable}
\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,inline]{Should the above table have pictures? If so, where? Since the table is already pretty dense}

@@

Instead, the first board was built out in a piecewise fashion, allowing each section of the board a way to fail without interrupting the rest of the board.
This allows each function of the board to be tested individually and in isolation so that changes could be made incrementally without needing to worry about their integration with the rest of the board.
\todoquestion{How do I convert this from first person to third person}For example, in order to build a system that would allow the board to be powered via USB or an internal battery, as well as charge the battery simultaneously, I would need to build a complex circuit consisting of parts neither I nor anyone I knew had used before.
Since power is obviously a critical component of the board, A secondary way of feeding power directly the board was built to bypass all that circuitry in case it fails.
In doing so, the board can still be powered and it's other features can be tested while a fix for that portion of the circuitry is developed for the next iteration.

@@

Finally, everything must connect to the CM4.
Figure \ref{fig:CM4Circuit} shows the the two mezzanine connectors the CM4 uses to communicate with the rest of the PCB (schematic in Figure \ref{fig:pcb_schematic_cm4}).
Remember that there is only 0.04mm of space in between each of the pins, leaving very little space to route the traces where they need to go.
The first thing to notice is the large stretch of traces on the back side of the board coming from the HDMI ports.
Because most of the traces come in positive and negative pairs, called differential pairs, the two tracks must be roughly the same length as each other.
This is because as data is transmitted across the pair of tracks, they need to arrive at their destination at the same time based on the physical distance the signals need to cover.
Since the speed of electricity is constant throughout the PCB board, the traces may need to wind back and forth so that the lengths of the positive and negative tracks are equal.
Also note the blank area surrounded by vias at the top of Figure \ref{fig:CM4Circuit}.
This is where the WiFi and Bluetooth antennas are located on the CM4 board.
By removing the copper pour from this area on the PCB board and stitching the perimeter with vias, it prevents the board from acting like another antenna and interfering with the signals being sent to and from the CM4.
Finally, each of the other components are connected to the CM4 board in the same manner as described previously in the chapter, such as the Ethernet adapter, the SD card reader, and remaining ports. See Appendices \ref{AppendixA} and \ref{AppendixB} for further information.

\begin{figure}[t]
  \centering
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=.75\linewidth]{Figures/kicad/close-ups/cm4-front}
    \caption{The front side of the PCB}
    \label{fig:CM4CircuitFront}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=.75\linewidth]{Figures/kicad/close-ups/cm4-back}
    \caption{The back side of the PCB}
    \label{fig:CM4CircuitBack}
  \end{subfigure}
  \caption{The circuitry surrounding the CM4's two mezzanine connectors}
  \label{fig:CM4Circuit}
\end{figure}

### 2022-03-29 paper/main.tex
Feedback from Advisor
Mostly just notes, they will be fixed next
Fix comments for Background chapter
Draft State of the Art Chapter
I need to pick up the pace
Start working on the Hardware chapter
Draft prototyping sections
The next section is going to be meaty, it might need to be broken up somehow
Appendix A and B
Designing the power, usb, hdmi and start cm4 circuits
Tweaks and updates to the hardware sections
Now a nice big sprint can begin. Just need the time

@@

\newcommand{\todosection}{\todo[linecolor=olive,backgroundcolor=olive!25,bordercolor=olive,inline]{TODO}}

@@

\thesistitle{Building a Mobile Device that Uses the Power of a Desktop Computer} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
\supervisor{Dr. Robert \textsc{Heinrichs}} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
\examiner{} % Your examiner's name, this is not currently used anywhere in the template, print it elsewhere with \examname
\degree{Software Engineering} % Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
\author{Dylan \textsc{Lathrum}} % Your name, this is used in the title page and abstract, print it elsewhere with \authorname
\addresses{} % Your address, this is not currently used anywhere in the template, print it elsewhere with \addressname

@@

		\vspace*{.06\textheight}
		{\scshape\LARGE \univname\par}\vspace{1.5cm} % University name
		\textsc{\Large Undergraduate Thesis}\\[0.5cm] % Thesis type

@@

\begin{abstract}
	\addchaptertocentry{\abstractname} % Add the abstract to the table of contents
	The Thesis Abstract is written here (and usually kept to just this page). The page is kept centered vertically so can expand into the blank space above the title too\ldots
\end{abstract}

@@

\begin{abbreviations}{ll} % Include a list of abbreviations (a table of two columns)

	\textbf{CPU} & \textbf{C}entral \textbf{P}rocessing \textbf{U}nit\\
	\textbf{CRD} & \textbf{C}hrome \textbf{R}emote \textbf{D}esktop\\
	\textbf{GPU} & \textbf{G}raphics \textbf{P}rocessing \textbf{U}nit\\
	\textbf{PC} & \textbf{P}ersonal \textbf{C}computer\\
	\textbf{PCB} & \textbf{P}rinted \textbf{C}ircuit \textbf{B}oard\\
	\textbf{RDP} & \textbf{R}emote \textbf{D}esktop \textbf{P}rotocol\\
	\textbf{SSH} & \textbf{S}ecure \textbf{Sh}ell (Protocol)\\
	\textbf{VNC} & \textbf{V}irtual \textbf{N}etwork \textbf{C}omputing\\

@@

\include{Appendices/AppendixA}
%\include{Appendices/AppendixB}
%\include{Appendices/AppendixC}

>>>>

\usepackage{wrapfig} % Required to wrap figures
\usepackage{subcaption} % Required for subfigures

@@

\setlength{\marginparwidth}{3cm} % The width of the margin for todo notes
\newcommand{\todoadvisor}[1]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue]{#1}}
\newcommand{\todoquestion}[1]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green]{#1}}
\newcommand{\todosection}{\todo[linecolor=olive,backgroundcolor=olive!25,bordercolor=olive,inline]{TODO}}

\usepackage{csvsimple} % Import CSV files for use in tables
\usepackage{longtable} % Tables that span multiple pages

@@

\thesistitle{Building a Mobile Device that Leverages the Power of a Desktop Computer} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
\supervisor{Dr. Robert \textsc{Heinrichs}} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
\examiner{} % Your examiner's name, this is not currently used anywhere in the template, print it elsewhere with \examname
\degree{Software Engineering} % Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
\author{Dylan \textsc{Lathrum}} % Your name, this is used in the title page and abstract, print it elsewhere with \authorname
\addresses{} % Your address, this is not currently used anywhere in the template, print it elsewhere with \addressname

@@

		\vspace*{.06\textheight}
		{\scshape\LARGE \univname\par}\vspace{1.5cm} % University name
		\textsc{\Large Honors Thesis}\\[0.5cm] % Thesis type

@@

\begin{abstract}
\addchaptertocentry{\abstractname} % Add the abstract to the table of contents

With the recent focus of attention towards remote work and mobile computing, the possibility of taking a powerful workstation wherever needed is enticing.
However, even emerging laptops today struggle to compete with desktops in terms of cost, maintenance, and future upgrades.
The price point of a powerful laptop is considerably higher compared to an equally powerful desktop computer, and most laptops are manufactured in a way that makes upgrading parts of the machine difficult or impossible, forcing a complete purchase in the event of failure or a component needing an upgrade.

In the case where someone already owns a desktop computer and must be mobile, instead of needing to purchase a second device at full price, it may be possible to develop a low-cost computer that has just enough power to connect to the existing desktop and run all processing there, using the mobile device only as a user interface.

\end{abstract}

@@

\begin{abbreviations}{ll} % Include a list of abbreviations (a table of two columns)
	\addchaptertocentry{\abbrevname} % Add the Abbreviations to the table of contents
	\textbf{CM4} & Raspberry Pi \textbf{C}ompute \textbf{M}odule \textbf{4}\\
	\textbf{CPU} & \textbf{C}entral \textbf{P}rocessing \textbf{U}nit\\
	\textbf{CRD} & \textbf{C}hrome \textbf{R}emote \textbf{D}esktop\\
	\textbf{FFC} & \textbf{F}lat \textbf{F}lexible \textbf{C}able\\
	\textbf{GPU} & \textbf{G}raphics \textbf{P}rocessing \textbf{U}nit\\
	\textbf{IDE} & \textbf{I}ntegrated \textbf{D}evelopment \textbf{E}nvironment\\
	\textbf{PC} & \textbf{P}ersonal \textbf{C}omputer\\
	\textbf{PCB} & \textbf{P}rinted \textbf{C}ircuit \textbf{B}oard\\
	\textbf{RDP} & \textbf{R}emote \textbf{D}esktop \textbf{P}rotocol\\
	\textbf{SBC} & \textbf{S}ingle \textbf{B}oard \textbf{C}omputer\\
	\textbf{SMD} & \textbf{S}urface \textbf{M}ounted \textbf{D}evice\\
	\textbf{SSH} & \textbf{S}ecure \textbf{Sh}ell (Protocol)\\
	\textbf{VNC} & \textbf{V}irtual \textbf{N}etwork \textbf{C}omputing\\

@@

\include{Appendices/AppendixA}
\include{Appendices/AppendixB}
%\include{Appendices/AppendixC}
